---
title: "R Notebook"
output: html_notebook
---

```{r setup}
rm(list=ls(all=TRUE))
    knitr::opts_knit$set(root.dir = normalizePath("/Users/paultran/Box/Students/T1D_complications")) 
```

Target journal: Diabetes Care IF 16
Clinical Risk of Type 1 Diabetes Complications: A tool to motivate change in patient behavior
Neuropathy, retinopathy, and nephropathy are common complications
patients seeing their true risk and working on modifiable risk factors are the best methods for primary prevention of T1D complications

We used an independent cross sectional study, PAGODA, to develop a clinical risk score for diabetic peripheral neuropathy in T1D patients. 

#Table 1. PAGODA cohort characteristics
n patients, sex, age distribution, duration distribution, DQ/DR genetics, PMH of cardiovascular diseases (HTN,MI,TIA,CVA), DPN, DAN, Retino, Nephro, Dyslipidemia. Distribution of cont variables, BP, all Lipids (cholesterol, LDL, HDL, triglycerides), BUN, Creatinine, and others

#Figure 1. Venn Diagram

```{r}
library(limma)
T1Dcomp_alldata<-read.csv("2_T1DComplications data_cleaned for Paul.csv")
attach(T1Dcomp_alldata)

T1Dcomp_alldata$DPN<-ifelse(Peripheral_Neuropathy3=="Yes",1,0)
T1Dcomp_alldata$CAN<-ifelse(Autonomic_Neuropathy3=="Yes",1,0)
T1Dcomp_alldata$DR<-ifelse(DM_Retinopathy3=="Yes",1,0)
T1Dcomp_alldata$DN<-ifelse(DM_Nephropathy3=="Yes",1,0)

attach(T1Dcomp_alldata)

a<-vennCounts(cbind(DPN,CAN,DR,DN))

table(Peripheral_Neuropathy3);table(Autonomic_Neuropathy3);table(DM_Retinopathy3);table(DM_Nephropathy3);table(Sex);table(Race);table(HTN3);table(Dyslipidemia3);table(Blindness_DM3);table(Photocoagulation3);table(CAD3);table(Prior_Angioplasty_Stent3);table(Prior_CABG3);table(Prior_CVA3);table(Prior_MI3);table(Prior_TIA3);table(CAD3);table(Amputation3);table(Diabetic_Foot_Ulcer3);table(Peripheral_Neuropathy_Type);table(Smoking); summary(Dr_Age);summary(Dx_Age);summary(Dur);summary(Hemoglobin);summary(Albumin);summary(LDL);summary(Total_Cholesterol);summary(Triglycerides);summary(HDL);summary(avg.BUN);summary(avg.Creatinine);summary(avg.Systolic);summary(avg.diastolic);summary(avg.MicroAlb);summary(AvgA1c);summary(ACR); vennDiagram(a)

```

Table 2. Univariate Predictors of T1D Complications
Table with OR 95% CI for each clinical variable and its association with the 4 complications


Univariate associations
```{r}
T1Dcomp_alldata<-read.csv("2_T1DComplications data_cleaned for Paul.csv")
attach(T1Dcomp_alldata)
T1Dcomp_alldata$DPN<-ifelse(Peripheral_Neuropathy3=="Yes",1,0)
T1Dcomp_alldata$CAN<-ifelse(Autonomic_Neuropathy3=="Yes",1,0)
T1Dcomp_alldata$DR<-ifelse(DM_Retinopathy3=="Yes",1,0)
T1Dcomp_alldata$DN<-ifelse(DM_Nephropathy3=="Yes",1,0)

vars <-
  c(
    "Peripheral_Neuropathy3",
    "Autonomic_Neuropathy3",
    "DM_Retinopathy3",
    "DM_Nephropathy3",
    
    "Blindness_DM3",
    "Photocoagulation3",
    "Amputation3",
    "Diabetic_Foot_Ulcer3",
    "Peripheral_Neuropathy_Type",
    
    "Sex",
    "Race",
    "Smoking",
    
    "HTN3",
    "Dyslipidemia3",
    "CAD3",
    "Prior_Angioplasty_Stent3",
    "Prior_CABG3",
    "Prior_CVA3",
    "Prior_MI3",
    "Prior_TIA3",
    
    "Dr_Age",
    "Dx_Age",
    "Dur",
    
    "Hemoglobin",
    "Albumin",
    "LDL",
    "Total_Cholesterol",
    "Triglycerides",
    "HDL",
    "avg.BUN",
    "avg.Creatinine",
    "avg.Systolic",
    "avg.diastolic",
    "avg.MicroAlb",
    "ACR",
    "AvgA1c",
    "SDA1c",
    "HbA1c_3.1",
    "Max_HbA1c",
    
    "Insulin..Aspart..Human",
    "Insulin..Lispro..Human",
    "insulin.glargine..rDNA.origin..injection",
    "acetYes",
    "atorvastatin.calcium",
    "levothYes",
    "lisinopril",
    "ThYes",
    "Simvastatin",
    "ramipril",
    "ibandronate.sodium",
    "Escitalopram",
    "Insulin..Isophane..Human",
    "Valsartan",
    "insulin.detemir",
    "insulin.glulisine.RDNA.origin",
    "Insulin.regular.human.recombinant",
    "insulin..human.recombinant",
    "lovastatin"
  )
suppressWarnings(
cbind(
  t(sapply(paste("DPN ~",vars),
                      function(form) {
  res.logist<-glm(form,data=T1Dcomp_alldata,family="binomial")
  summary(res.logist)$coefficients[2,]
  }
  )),
  t(sapply(paste("CAN ~",vars),
                      function(form) {
  res.logist<-glm(form,data=T1Dcomp_alldata,family="binomial")
  summary(res.logist)$coefficients[2,]
  }
  )),
  t(sapply(paste("DR ~",vars),
                      function(form) {
  res.logist<-glm(form,data=T1Dcomp_alldata,family="binomial")
  summary(res.logist)$coefficients[2,]
  }
  )),
  t(sapply(paste("DN ~",vars),
                      function(form) {
  res.logist<-glm(form,data=T1Dcomp_alldata,family="binomial")
  summary(res.logist)$coefficients[2,]
  }
  ))
)
)


```


Figure 2. Multivariate logistic regression models for 4 complications
A. Variables with coefficients
B. Nomogram
C. AUC of test data
D. Interation of Age and T1D Dur

Finding is that clinical variables are good predictors of T1D complication risk. Interaction is interesting

how to optimize model? rfe, manual selection, lasso
which variables to start with? 
DPN: Autonomic_Neuropathy3+DM_Retinopathy3+DM_Nephropathy3+Sex+HTN+avg.Systolic+avg.diastolic+Dyslipidemia+CAD3+Dr_Age+Dx_Age+Dur+Hemoglobin+avg.BUN
CAN:


#interaction of age and dur

```{r}

library(caret)
library(doParallel)
library(dplyr)
library(pROC)

T1Dcomp_alldata<-read.csv("2_T1DComplications data_cleaned for Paul.csv")
attach(T1Dcomp_alldata)
T1Dcomp_alldata$DPN<-ifelse(Peripheral_Neuropathy3=="Yes",1,0)
T1Dcomp_alldata$CAN<-ifelse(Autonomic_Neuropathy3=="Yes",1,0)
T1Dcomp_alldata$DR<-ifelse(DM_Retinopathy3=="Yes",1,0)
T1Dcomp_alldata$DN<-ifelse(DM_Nephropathy3=="Yes",1,0)



 # 1) DPN
dpn_idx = createDataPartition(Peripheral_Neuropathy3, p = 0.75, list = FALSE)
dpn_trn = T1Dcomp_alldata[dpn_idx, ] #548
dpn_tst = T1Dcomp_alldata[-dpn_idx, ] #182

set.seed(123)
my_dpn_int_model <- train(
  form = Peripheral_Neuropathy3~Dr_Age*Dur,
  method = "glm",
  family = "binomial",
  data = dpn_trn,
  trControl = trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 5),
  na.action = na.pass
)

my_dpn_add_model <- train(
  form = Peripheral_Neuropathy3~Dr_Age+Dur,
  method = "glm",
  family = "binomial",
  data = dpn_trn,
  trControl = trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 5),
  na.action = na.pass
)


dpn.aov<-anova(my_dpn_add_model$finalModel,my_dpn_int_model$finalModel)

my_dpn_add_model$results;my_dpn_int_model$results
results2<-resamples(list(mod1=my_dpn_add_model,mod2=my_dpn_int_model))
summary(results2);bwplot(results2);dpn.aov

dpn_test_prob = predict(my_dpn_add_model, newdata = dpn_tst, type = "prob",na.action = na.pass)
dpn_test_roc = roc(dpn_tst$Peripheral_Neuropathy3 ~ dpn_test_prob$Yes, plot = TRUE, print.auc = TRUE)


 # 2) DAN
dan_idx = createDataPartition(Autonomic_Neuropathy3, p = 0.75, list = FALSE)
dan_trn = T1Dcomp_alldata[dan_idx, ] #548
dan_tst = T1Dcomp_alldata[-dan_idx, ] #182

set.seed(123)
my_dan_add_model <- train(
  form = Autonomic_Neuropathy3~Dr_Age+Dur,
  method = "glm",
  family = "binomial",
  data = dan_trn,
  trControl = trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 5),
  na.action = na.pass
)

my_dan_int_model <- train(
  form = Autonomic_Neuropathy3~Dr_Age*Dur,
  method = "glm",
  family = "binomial",
  data = dan_trn,
  trControl = trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 5),
  na.action = na.pass
)

dan.aov<-anova(my_dan_add_model$finalModel,my_dan_int_model$finalModel)
results2<-resamples(list(mod1=my_dan_add_model,mod2=my_dan_int_model))
summary(results2);bwplot(results2);dan.aov;summary(my_dan_add_model);summary(my_dan_int_model)


dan_test_prob = predict(my_dan_add_model, newdata = dan_tst, type = "prob",na.action = na.pass)
dan_test_roc = roc(dan_tst$Autonomic_Neuropathy3 ~ dan_test_prob$Yes, plot = TRUE, print.auc = TRUE)


 # 3) Dret

dret_idx = createDataPartition(neuro_dat_formodel$class, p = 0.75, list = FALSE)
dret_trn = neuro_dat_formodel[dret_idx, ] #548
dret_tst = neuro_dat_formodel[-dret_idx, ] #182

set.seed(123)
my_dret_model <- train(
  form = class~HTN3+Dyslipidemia3+Dr_Age*Dur+avg.Systolic+AvgA1c+Peripheral_Neuropathy2020,
  method = "glm",
  family = "binomial",
  data = dret_trn,
  trControl = trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 5),
  na.action = na.pass
)

dret_test_prob = predict(my_dret_model, newdata = dret_tst, type = "prob",na.action = na.pass)
dret_test_roc = roc(dret_tst$class ~ dret_test_prob$Yes, plot = TRUE, print.auc = TRUE)

 # 4) Dnep
neuro_dat_formodel$class<-recode_factor(neuro_alldata$Nephropathy,
              `0` = "No",
              `1` = "Yes")

dnep_idx = createDataPartition(neuro_dat_formodel$class, p = 0.75, list = FALSE)
dnep_trn = neuro_dat_formodel[dnep_idx, ] #548
dnep_tst = neuro_dat_formodel[-dnep_idx, ] #182

set.seed(123)
my_dnep_model <- train(
  form = class~HTN3+Dyslipidemia3+Dr_Age*Dur+avg.Systolic+AvgA1c+Peripheral_Neuropathy2020,
  method = "glm",
  family = "binomial",
  data = dnep_trn,
  trControl = trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 5),
  na.action = na.pass
)

dnep_test_prob = predict(my_dnep_model, newdata = dnep_tst, type = "prob",na.action = na.pass)
dnep_test_roc = roc(dnep_tst$class ~ dnep_test_prob$Yes, plot = TRUE, print.auc = TRUE)


#Summary data
summary(my_dpn_model);summary(my_dan_model);summary(my_dret_model);summary(my_dnep_model);dpn_test_roc;dan_test_roc;dret_test_roc;dnep_test_roc

```





#Comparing predictive models
```{r}

library(caret)
library(doParallel)
library(dplyr)
library(pROC)
library(glmnet)
source("my_functions.R")
T1Dcomp_alldata<-read.csv("2_T1DComplications data_cleaned for Paul.csv")

set.seed(123)



 # 1) DPN

dpn <- suppressWarnings(compare_3_models(T1Dcomp_alldata,"Peripheral_Neuropathy3",varstoassess = c("Peripheral_Neuropathy3","Autonomic_Neuropathy3","DM_Retinopathy3","DM_Nephropathy3","HTN3","Dyslipidemia3","CAD3","Prior_TIA3","Prior_MI3","Prior_CVA3","Dr_Age","Dur","Hemoglobin","Albumin","avg.Creatinine","avg.Systolic","avg.BUN","avg.MicroAlb","AvgA1c","Sex","HDL","Max_HbA1c")))

 # 2) DAN
can <- compare_3_models(T1Dcomp_alldata,"Autonomic_Neuropathy3",varstoassess = c("Autonomic_Neuropathy3","Peripheral_Neuropathy3","DM_Retinopathy3","DM_Nephropathy3","HTN3","Dyslipidemia3","CAD3","Prior_TIA3","Prior_MI3","Prior_CVA3","Dr_Age","Dur","Hemoglobin","Albumin","avg.Creatinine","avg.Systolic","avg.BUN","avg.MicroAlb","AvgA1c","Sex","HDL","Max_HbA1c"))

 # 3) Dret
dr <- compare_3_models(T1Dcomp_alldata,"DM_Retinopathy3",varstoassess = c("DM_Retinopathy3","Peripheral_Neuropathy3","Autonomic_Neuropathy3","DM_Nephropathy3","HTN3","Dyslipidemia3","CAD3","Prior_TIA3","Prior_MI3","Prior_CVA3","Dr_Age","Dur","Hemoglobin","Albumin","avg.Creatinine","avg.Systolic","avg.BUN","avg.MicroAlb","AvgA1c","Sex","HDL","Max_HbA1c"))

 # 4) Dnep
dn <- compare_3_models(T1Dcomp_alldata,"DM_Nephropathy3",varstoassess = c("DM_Nephropathy3","Peripheral_Neuropathy3","Autonomic_Neuropathy3","DM_Retinopathy3","HTN3","Dyslipidemia3","CAD3","Prior_TIA3","Prior_MI3","Prior_CVA3","Dr_Age","Dur","Hemoglobin","Albumin","avg.Creatinine","avg.Systolic","avg.BUN","avg.MicroAlb","AvgA1c","Sex","HDL","Max_HbA1c"))


#Summary data
paste("____________________________DPN____________________________");dpn$lasso_tst_roc$auc;dpn$rfelogit_tst_roc$auc;dpn$customlogit_tst_roc$auc;dpn$lasso_coef;dpn$rfelogit_coef;dpn$customlogit_coef;paste("____________________________CAN____________________________");can$lasso_tst_roc$auc;can$rfelogit_tst_roc$auc;can$customlogit_tst_roc$auc;can$lasso_coef;can$rfelogit_coef;can$customlogit_coef;paste("____________________________DR____________________________");dr$lasso_tst_roc$auc;dr$rfelogit_tst_roc$auc;dr$customlogit_tst_roc$auc;dr$lasso_coef;dr$rfelogit_coef;dr$customlogit_coef;paste("____________________________DN____________________________");dn$lasso_tst_roc$auc;dn$rfelogit_tst_roc$auc;dn$customlogit_tst_roc$auc;dn$lasso_coef;dn$rfelogit_coef;dn$customlogit_coef


```


#Nomogram
```{r}
#nomogram
library(rms)
T1Dcomp_alldata<-read.csv("2_T1DComplications data_cleaned for Paul.csv")


options(datadist='ddist')
ddist <-datadist(T1Dcomp_alldata)

# 1) dpn
dpn.mod<-lrm(Peripheral_Neuropathy3~Dr_Age+Dur+avg.Systolic+AvgA1c,data=T1Dcomp_alldata)
nom.dpn<-nomogram(dpn.mod,
                 lp.at = seq(-3,4,by=0.5),
                 fun = function(x)1/(1+exp(-x)),
                     fun.at=c(0.001,0.01,.05,seq(0.1,0.9,by=0.1),0.95,0.99,0.999),
                     funlabel="Risk of Complication",
                 conf.int=c(0.1,0.7),
                 abbrev=T,
                 minlength=1,lp=F)

# 2) dan
dan.mod<-lrm(Autonomic_Neuropathy3~Dr_Age+Dur+avg.Systolic+AvgA1c,data=T1Dcomp_alldata)
nom.dan<-nomogram(dan.mod,
                 lp.at = seq(-3,4,by=0.5),
                 fun = function(x)1/(1+exp(-x)),
                     fun.at=c(0.001,0.01,.05,seq(0.1,0.9,by=0.1),0.95,0.99,0.999),
                     funlabel="Risk of Complication",
                 conf.int=c(0.1,0.7),
                 abbrev=T,
                 minlength=1,lp=F)

# 3) dret
dret.mod<-lrm(DM_Retinopathy3~Dr_Age+Dur+avg.Systolic+AvgA1c,data=T1Dcomp_alldata)
nom.dret<-nomogram(dret.mod,
                 lp.at = seq(-3,4,by=0.5),
                 fun = function(x)1/(1+exp(-x)),
                     fun.at=c(0.001,0.01,.05,seq(0.1,0.9,by=0.1),0.95,0.99,0.999),
                     funlabel="Risk of Complication",
                 conf.int=c(0.1,0.7),
                 abbrev=T,
                 minlength=1,lp=F)

# 4) dnep
dnep.mod<-lrm(DM_Nephropathy3~Dr_Age+Dur+avg.Systolic+AvgA1c,data=T1Dcomp_alldata)
nom.dnep<-nomogram(dnep.mod,
                 lp.at = seq(-3,4,by=0.5),
                 fun = function(x)1/(1+exp(-x)),
                     fun.at=c(0.001,0.01,.05,seq(0.1,0.9,by=0.1),0.95,0.99,0.999),
                     funlabel="Risk of Complication",
                 conf.int=c(0.1,0.7),
                 abbrev=T,
                 minlength=1,lp=F)

#save data
saveRDS(dpn.mod,file = "/Users/paultran/Box/Students/T1D_complications/Shiny App/T1D_Complications/data/T1Dcomp_DPNmod_logit.rds")
saveRDS(dan.mod,file = "/Users/paultran/Box/Students/T1D_complications/Shiny App/T1D_Complications/data/T1Dcomp_DANmod_logit.rds")
saveRDS(dret.mod,file = "/Users/paultran/Box/Students/T1D_complications/Shiny App/T1D_Complications/data/T1Dcomp_DRETmod_logit.rds")
saveRDS(dnep.mod,file = "/Users/paultran/Box/Students/T1D_complications/Shiny App/T1D_Complications/data/T1Dcomp_DNEPmod_logit.rds")

#summary and plot
summary(dpn.mod);summary(dan.mod);summary(dret.mod);summary(dnep.mod);plot(nom.dpn);plot(nom.dan);plot(nom.dret);plot(nom.dnep)
```


Figure 3. Website developed to motivate change

Key features are risk reporting for 4 complications. How would risk be modified with changes in SBP or HbA1C? How patient compares to others of same Age and T1D dur for modifiable risk factors? Methods to modify risk factors (diet, weight loss association with BP). 

Conclusion: SBP and HbA1C are modifiable risk factors for T1D complications through lifestyle changes and medical management. Motivation to manage SBP and HbA1C may be more difficult for younger individuals and T1D complication risk tool may aid with this.


give percentile for htn, a1c, and overall risk?

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

