---
title: "R Notebook"
output: html_notebook
---

```{r setup}
rm(list=ls(all=TRUE))
    knitr::opts_knit$set(root.dir = normalizePath("/Users/paultran/Box/Students/T1D_complications")) 
```

Clinical Risk of Type 1 Diabetes Complications: A tool to motivate change in patient behavior
Neuropathy, retinopathy, and nephropathy are common complications
patients seeing their true risk and working on modifiable risk factors are the best methods for primary prevention of T1D complications

We used an independent cross sectional study, PAGODA, to develop a clinical risk score for diabetic peripheral neuropathy in T1D patients. 

#Table 1. PAGODA cohort characteristics

```{r}
library(limma)
T1Dcomp_alldata<-read.csv("2_T1DComplications data_cleaned for Paul.csv")
attach(T1Dcomp_alldata)

T1Dcomp_alldata$DPN<-ifelse(Peripheral_Neuropathy3=="Yes",1,0)
T1Dcomp_alldata$CAN<-ifelse(Autonomic_Neuropathy3=="Yes",1,0)
T1Dcomp_alldata$DR<-ifelse(DM_Retinopathy3=="Yes",1,0)
T1Dcomp_alldata$DN<-ifelse(DM_Nephropathy3=="Yes",1,0)

attach(T1Dcomp_alldata)

a<-vennCounts(cbind(DPN,CAN,DR,DN))

table(Peripheral_Neuropathy3);table(Autonomic_Neuropathy3);table(DM_Retinopathy3);table(DM_Nephropathy3);table(Sex);table(Race);table(HTN3);table(Dyslipidemia3);table(Blindness_DM3);table(Photocoagulation3);table(CAD3);table(Prior_Angioplasty_Stent3);table(Prior_CABG3);table(Prior_CVA3);table(Prior_MI3);table(Prior_TIA3);table(CAD3);table(Amputation3);table(Diabetic_Foot_Ulcer3);table(Peripheral_Neuropathy_Type);table(Smoking); summary(Dr_Age);summary(Dx_Age);summary(Dur);summary(Hemoglobin);summary(Albumin);summary(LDL);summary(Total_Cholesterol);summary(Triglycerides);summary(HDL);summary(avg.BUN);summary(avg.Creatinine);summary(avg.Systolic);summary(avg.diastolic);summary(avg.MicroAlb);summary(AvgA1c);summary(ACR); vennDiagram(a)

```

Table 2. Univariate Predictors of T1D Complications
Table with OR 95% CI for each clinical variable and its association with the 4 complications


Univariate associations
```{r}
T1Dcomp_alldata<-read.csv("2_T1DComplications data_cleaned for Paul.csv")
attach(T1Dcomp_alldata)
T1Dcomp_alldata$DPN<-ifelse(Peripheral_Neuropathy3=="Yes",1,0)
T1Dcomp_alldata$CAN<-ifelse(Autonomic_Neuropathy3=="Yes",1,0)
T1Dcomp_alldata$DR<-ifelse(DM_Retinopathy3=="Yes",1,0)
T1Dcomp_alldata$DN<-ifelse(DM_Nephropathy3=="Yes",1,0)

vars <-
  c(
    "Peripheral_Neuropathy3",
    "Autonomic_Neuropathy3",
    "DM_Retinopathy3",
    "DM_Nephropathy3",
    
    "Blindness_DM3",
    "Photocoagulation3",
    "Amputation3",
    "Diabetic_Foot_Ulcer3",
    "Peripheral_Neuropathy_Type",
    
    "Sex",
    "Race",
    "Smoking",
    
    "HTN3",
    "Dyslipidemia3",
    "CAD3",
    "Prior_Angioplasty_Stent3",
    "Prior_CABG3",
    "Prior_CVA3",
    "Prior_MI3",
    "Prior_TIA3",
    
    "Dr_Age",
    "Dx_Age",
    "Dur",
    
    "Hemoglobin",
    "Albumin",
    "LDL",
    "Total_Cholesterol",
    "Triglycerides",
    "HDL",
    "avg.BUN",
    "avg.Creatinine",
    "avg.Systolic",
    "avg.diastolic",
    "avg.MicroAlb",
    "ACR",
    "AvgA1c",
    "SDA1c",
    "HbA1c_3.1",
    "Max_HbA1c",
    
    "Insulin..Aspart..Human",
    "Insulin..Lispro..Human",
    "insulin.glargine..rDNA.origin..injection",
    "acetYes",
    "atorvastatin.calcium",
    "levothYes",
    "lisinopril",
    "ThYes",
    "Simvastatin",
    "ramipril",
    "ibandronate.sodium",
    "Escitalopram",
    "Insulin..Isophane..Human",
    "Valsartan",
    "insulin.detemir",
    "insulin.glulisine.RDNA.origin",
    "Insulin.regular.human.recombinant",
    "insulin..human.recombinant",
    "lovastatin"
  )
suppressWarnings(
cbind(
  t(sapply(paste("DPN ~",vars),
                      function(form) {
  res.logist<-glm(form,data=T1Dcomp_alldata,family="binomial")
  summary(res.logist)$coefficients[2,]
  }
  )),
  t(sapply(paste("CAN ~",vars),
                      function(form) {
  res.logist<-glm(form,data=T1Dcomp_alldata,family="binomial")
  summary(res.logist)$coefficients[2,]
  }
  )),
  t(sapply(paste("DR ~",vars),
                      function(form) {
  res.logist<-glm(form,data=T1Dcomp_alldata,family="binomial")
  summary(res.logist)$coefficients[2,]
  }
  )),
  t(sapply(paste("DN ~",vars),
                      function(form) {
  res.logist<-glm(form,data=T1Dcomp_alldata,family="binomial")
  summary(res.logist)$coefficients[2,]
  }
  ))
)
)


```

after reading more into harrellverse stuff, have ideas of things to try
mostly inspired by this blog: https://www.nicholas-ollberding.com/post/an-introduction-to-the-harrell-verse-predictive-modeling-using-the-hmisc-and-rms-packages/

use spiked histogram to see if something happens around puberty onset t1d
use spiked histogram to assess nonlinearity
use rcs for age and dur
try out the fancy interaction functions and plots
predict multiple bp and hba1cs on one plot
chunk test the related variables

read this for ideas about the proteins, something about rcs and pca
https://www.fharrell.com/post/score-data-reduction/

outline of analysis
construct a full model of all possible clinical variables
compare to subset with 4 variables
cant fit rcs for all continuous because too many df for variables


```{r}
#code ideas from reading harrellverse stuff
library(ggplot2)
library(rms)
#load and subset data
vars <-
  c(
    "Peripheral_Neuropathy3",
    "Autonomic_Neuropathy3",
    "DM_Retinopathy3",
    "DM_Nephropathy3",
    
    "Blindness_DM3",
    "Photocoagulation3",
    "Amputation3",
    "Diabetic_Foot_Ulcer3",
    "Peripheral_Neuropathy_Type",
    
    "Sex",
    "Race",
    "Smoking",
    
    "HTN3",
    "Dyslipidemia3",
    "CAD3",
    "Prior_Angioplasty_Stent3",
    "Prior_CABG3",
    "Prior_CVA3",
    "Prior_MI3",
    "Prior_TIA3",
    
    "Dr_Age",
    "Dx_Age",
    "Dur",
    
    "Hemoglobin",
    "Albumin",
    "LDL",
    "Total_Cholesterol",
    "Triglycerides",
    "HDL",
    "avg.BUN",
    "avg.Creatinine",
    "avg.Systolic",
    "avg.diastolic",
    "avg.MicroAlb",
    "ACR",
    "AvgA1c",
    "SDA1c",
    "HbA1c_3.1",
    "Max_HbA1c",
    
    "Insulin..Aspart..Human",
    "Insulin..Lispro..Human",
    "insulin.glargine..rDNA.origin..injection",
    "acetYes",
    "atorvastatin.calcium",
    "levothYes",
    "lisinopril",
    "ThYes",
    "Simvastatin",
    "ramipril",
    "ibandronate.sodium",
    "Escitalopram",
    "Insulin..Isophane..Human",
    "Valsartan",
    "insulin.detemir",
    "insulin.glulisine.RDNA.origin",
    "Insulin.regular.human.recombinant",
    "insulin..human.recombinant",
    "lovastatin"
  )
T1Dcomp_alldata<-read.csv("2_T1DComplications data_cleaned for Paul.csv")[,vars]
T1Dcomp_alldata$Peripheral_Neuropathy3<-ifelse(T1Dcomp_alldata$Peripheral_Neuropathy3=="Yes",1,0)
T1Dcomp_alldata$Autonomic_Neuropathy3<-ifelse(T1Dcomp_alldata$Autonomic_Neuropathy3=="Yes",1,0)
T1Dcomp_alldata$DM_Retinopathy3<-ifelse(T1Dcomp_alldata$DM_Retinopathy3=="Yes",1,0)
T1Dcomp_alldata$DM_Nephropathy3<-ifelse(T1Dcomp_alldata$DM_Nephropathy3=="Yes",1,0)
dd <- datadist(T1Dcomp_alldata)
options(datadist = "dd")

#describe all variables
# Hmisc::describe(T1Dcomp_alldata)

#summary table stratified by outcome
# s <- Hmisc::summaryM(    Autonomic_Neuropathy3+
#     DM_Retinopathy3+
#     DM_Nephropathy3+
#     
#     Blindness_DM3+
#     Photocoagulation3+
#     Amputation3+
#     Diabetic_Foot_Ulcer3+
#     Peripheral_Neuropathy_Type+
#     
#     Sex+
#     Race+
#     Smoking+
#     
#     HTN3+
#     Dyslipidemia3+
#     CAD3+
#     Prior_Angioplasty_Stent3+
#     Prior_CABG3+
#     Prior_CVA3+
#     Prior_MI3+
#     Prior_TIA3+
#     
#     Dr_Age+
#     Dx_Age+
#     Dur+
#     
#     Hemoglobin+
#     Albumin+
#     LDL+
#     Total_Cholesterol+
#     Triglycerides+
#     HDL+
#     avg.BUN+
#     avg.Creatinine+
#     avg.Systolic+
#     avg.diastolic+
#     avg.MicroAlb+
#     ACR+
#     AvgA1c+
#     SDA1c+
#     HbA1c_3.1+
#     Max_HbA1c+
#     
#     Insulin..Aspart..Human+
#     Insulin..Lispro..Human+
#     insulin.glargine..rDNA.origin..injection+
#     acetYes+
#     atorvastatin.calcium+
#     levothYes+
#     lisinopril+
#     ThYes+
#     Simvastatin+
#     ramipril+
#     ibandronate.sodium+
#     Escitalopram+
#     Insulin..Isophane..Human+
#     Valsartan+
#     insulin.detemir+
#     insulin.glulisine.RDNA.origin+
#     Insulin.regular.human.recombinant+
#     insulin..human.recombinant+
#     lovastatin  ~ Peripheral_Neuropathy3, data = T1Dcomp_alldata,
# overall = TRUE, test = TRUE, continuous = 5)
# 
# s

#spiked histogram to see distribution of risk
a <- ggplot(T1Dcomp_alldata, aes(x = Dr_Age, y = Peripheral_Neuropathy3)) +
  Hmisc::histSpikeg(Peripheral_Neuropathy3 ~ Dr_Age, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nAge", y = "Probability(DPN)\n")
b <- ggplot(T1Dcomp_alldata, aes(x = Dr_Age, y = Autonomic_Neuropathy3)) +
  Hmisc::histSpikeg(Autonomic_Neuropathy3 ~ Dr_Age, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nAge", y = "Probability(AN)\n")
c <- ggplot(T1Dcomp_alldata, aes(x = Dr_Age, y = DM_Retinopathy3)) +
  Hmisc::histSpikeg(DM_Retinopathy3 ~ Dr_Age, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nAge", y = "Probability(DR)\n")
d <- ggplot(T1Dcomp_alldata, aes(x = Dr_Age, y = DM_Nephropathy3)) +
  Hmisc::histSpikeg(DM_Nephropathy3 ~ Dr_Age, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nAge", y = "Probability(DN)\n")

e <- ggplot(T1Dcomp_alldata, aes(x = Dx_Age, y = Peripheral_Neuropathy3)) +
  Hmisc::histSpikeg(Peripheral_Neuropathy3 ~ Dx_Age, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nDx_Age", y = "Probability(DPN)\n")
f <- ggplot(T1Dcomp_alldata, aes(x = Dx_Age, y = Autonomic_Neuropathy3)) +
  Hmisc::histSpikeg(Autonomic_Neuropathy3 ~ Dx_Age, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nDx_Age", y = "Probability(AN)\n")
g <- ggplot(T1Dcomp_alldata, aes(x = Dx_Age, y = DM_Retinopathy3)) +
  Hmisc::histSpikeg(DM_Retinopathy3 ~ Dx_Age, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nDx_Age", y = "Probability(DR)\n")
h <- ggplot(T1Dcomp_alldata, aes(x = Dx_Age, y = DM_Nephropathy3)) +
  Hmisc::histSpikeg(DM_Nephropathy3 ~ Dx_Age, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nDx_Age", y = "Probability(DN)\n")

i <- ggplot(T1Dcomp_alldata, aes(x = Dur, y = Peripheral_Neuropathy3)) +
  Hmisc::histSpikeg(Peripheral_Neuropathy3 ~ Dur, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nT1D_Dur", y = "Probability(DPN)\n")
j <- ggplot(T1Dcomp_alldata, aes(x = Dur, y = Autonomic_Neuropathy3)) +
  Hmisc::histSpikeg(Autonomic_Neuropathy3 ~ Dur, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nT1D_Dur", y = "Probability(AN)\n")
k <- ggplot(T1Dcomp_alldata, aes(x = Dur, y = DM_Retinopathy3)) +
  Hmisc::histSpikeg(DM_Retinopathy3 ~ Dur, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nT1D_Dur", y = "Probability(DR)\n")
l <- ggplot(T1Dcomp_alldata, aes(x = Dur, y = DM_Nephropathy3)) +
  Hmisc::histSpikeg(DM_Nephropathy3 ~ Dur, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nT1D_Dur", y = "Probability(DN)\n")

m <- ggplot(T1Dcomp_alldata, aes(x = AvgA1c, y = Peripheral_Neuropathy3)) +
  Hmisc::histSpikeg(Peripheral_Neuropathy3 ~ AvgA1c, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nAvgA1c", y = "Probability(DPN)\n")
n <- ggplot(T1Dcomp_alldata, aes(x = AvgA1c, y = Autonomic_Neuropathy3)) +
  Hmisc::histSpikeg(Autonomic_Neuropathy3 ~ AvgA1c, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nAvgA1c", y = "Probability(AN)\n")
o <- ggplot(T1Dcomp_alldata, aes(x = AvgA1c, y = DM_Retinopathy3)) +
  Hmisc::histSpikeg(DM_Retinopathy3 ~ AvgA1c, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nAvgA1c", y = "Probability(DR)\n")
p <- ggplot(T1Dcomp_alldata, aes(x = AvgA1c, y = DM_Nephropathy3)) +
  Hmisc::histSpikeg(DM_Nephropathy3 ~ AvgA1c, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nAvgA1c", y = "Probability(DN)\n")

q <- ggplot(T1Dcomp_alldata, aes(x = avg.Systolic, y = Peripheral_Neuropathy3)) +
  Hmisc::histSpikeg(Peripheral_Neuropathy3 ~ avg.Systolic, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nSystolic", y = "Probability(DPN)\n")
r <- ggplot(T1Dcomp_alldata, aes(x = avg.Systolic, y = Autonomic_Neuropathy3)) +
  Hmisc::histSpikeg(Autonomic_Neuropathy3 ~ avg.Systolic, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nSystolic", y = "Probability(AN)\n")
s <- ggplot(T1Dcomp_alldata, aes(x = avg.Systolic, y = DM_Retinopathy3)) +
  Hmisc::histSpikeg(DM_Retinopathy3 ~ avg.Systolic, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nSystolic", y = "Probability(DR)\n")
t <- ggplot(T1Dcomp_alldata, aes(x = avg.Systolic, y = DM_Nephropathy3)) +
  Hmisc::histSpikeg(DM_Nephropathy3 ~ avg.Systolic, lowess = TRUE, data = T1Dcomp_alldata) +
  labs(x = "\nSystolic", y = "Probability(DN)\n")

 #use rcs
DPNselect <- lrm(Peripheral_Neuropathy3 ~
    rcs(Dr_Age,3)+
    Dur+
    avg.Systolic+
    rcs(Dx_Age,5)+
    rcs(AvgA1c,4),
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)
ANselect <- lrm(Autonomic_Neuropathy3 ~
    rcs(Dr_Age,3)+
    Dur+
    rcs(Dx_Age,5)+
    avg.Systolic+
    rcs(AvgA1c,4),
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)
DRselect <- lrm(DM_Retinopathy3 ~
    rcs(Dr_Age,3)+
    rcs(Dur,3)+
    rcs(Dx_Age,5)+
    avg.Systolic+
    rcs(AvgA1c,4),
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)
DNselect <- lrm(DM_Nephropathy3 ~
    rcs(Dr_Age,3)+
    Dur+
    rcs(Dx_Age,5)+
    avg.Systolic+
    rcs(AvgA1c,4),
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)

bDPNselect <- lrm(Peripheral_Neuropathy3 ~
    rcs(Dr_Age,3)+
    avg.Systolic+
    rcs(Dx_Age,5)+
    rcs(AvgA1c,4),
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)
bANselect <- lrm(Autonomic_Neuropathy3 ~
    rcs(Dr_Age,3)+
    rcs(Dx_Age,5)+
    avg.Systolic+
    rcs(AvgA1c,4),
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)
bDRselect <- lrm(DM_Retinopathy3 ~
    rcs(Dr_Age,3)+
    rcs(Dx_Age,5)+
    avg.Systolic+
    rcs(AvgA1c,4),
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)
bDNselect <- lrm(DM_Nephropathy3 ~
    rcs(Dr_Age,3)+
    rcs(Dx_Age,5)+
    avg.Systolic+
    rcs(AvgA1c,4),
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)


#multiple values of any variable
a2<-ggplot(Predict(bDPNselect, Dr_Age, AvgA1c = c(5, 7, 9,11), fun = plogis))
c2<-ggplot(Predict(bDPNselect, Dr_Age, avg.Systolic = c(120, 140, 160, 180), fun = plogis))

e2<-ggplot(Predict(bANselect, Dr_Age, AvgA1c = c(5, 7, 9,11), fun = plogis))
g2<-ggplot(Predict(bANselect, Dr_Age, avg.Systolic = c(120, 140, 160, 180), fun = plogis))

i2<-ggplot(Predict(DRselect, Dr_Age, AvgA1c = c(5, 7, 9,11), fun = plogis))
k2<-ggplot(Predict(DRselect, Dr_Age, avg.Systolic = c(120, 140, 160, 180), fun = plogis))

m2<-ggplot(Predict(bDNselect, Dr_Age, AvgA1c = c(5, 7, 9,11), fun = plogis))
o2<-ggplot(Predict(bDNselect, Dr_Age, avg.Systolic = c(120, 140, 160, 180), fun = plogis))

#nomogram
nom.dpn<-nomogram(bDPNselect,
                 lp.at = seq(-3,4,by=0.5),
                 fun = function(x)1/(1+exp(-x)),
                     fun.at=c(0.001,0.01,.05,seq(0.1,0.9,by=0.1),0.95,0.99,0.999),
                     funlabel="Risk of DPN",
                 conf.int=c(0.1,0.7),
                 abbrev=T,
                 minlength=1,lp=F)
nom.an<-nomogram(bANselect,
                 lp.at = seq(-3,4,by=0.5),
                 fun = function(x)1/(1+exp(-x)),
                     fun.at=c(0.001,0.01,.05,seq(0.1,0.9,by=0.1),0.95,0.99,0.999),
                     funlabel="Risk of AN",
                 conf.int=c(0.1,0.7),
                 abbrev=T,
                 minlength=1,lp=F)
nom.dr<-nomogram(DRselect,
                 lp.at = seq(-3,4,by=0.5),
                 fun = function(x)1/(1+exp(-x)),
                     fun.at=c(0.001,0.01,.05,seq(0.1,0.9,by=0.1),0.95,0.99,0.999),
                     funlabel="Risk of DR",
                 conf.int=c(0.1,0.7),
                 abbrev=T,
                 minlength=1,lp=F)
nom.dn<-nomogram(bDNselect,
                 lp.at = seq(-3,4,by=0.5),
                 fun = function(x)1/(1+exp(-x)),
                     fun.at=c(0.001,0.01,.05,seq(0.1,0.9,by=0.1),0.95,0.99,0.999),
                     funlabel="Risk of DN",
                 conf.int=c(0.1,0.7),
                 abbrev=T,
                 minlength=1,lp=F)

#save data
saveRDS(bDPNselect,file = "/Users/paultran/Box/Students/T1D_complications/Shiny App/T1D_Complications/data/T1Dcomp_DPNmod_logit.rds")
saveRDS(bANselect,file = "/Users/paultran/Box/Students/T1D_complications/Shiny App/T1D_Complications/data/T1Dcomp_DANmod_logit.rds")
saveRDS(DRselect,file = "/Users/paultran/Box/Students/T1D_complications/Shiny App/T1D_Complications/data/T1Dcomp_DRETmod_logit.rds")
saveRDS(bDNselect,file = "/Users/paultran/Box/Students/T1D_complications/Shiny App/T1D_Complications/data/T1Dcomp_DNEPmod_logit.rds")

#summary and plot
lrtest(DPNselect,bDPNselect);lrtest(ANselect,bANselect);lrtest(DRselect,bDRselect);lrtest(DNselect,bDNselect);summary(bDPNselect);summary(bANselect);summary(DRselect);summary(bDNselect);plot(anova(bDPNselect), what='proportion chisq');plot(anova(bANselect), what='proportion chisq');plot(anova(DRselect), what='proportion chisq');plot(anova(bDNselect), what='proportion chisq');plot(Predict(bDPNselect, fun=plogis));plot(Predict(bANselect, fun=plogis));plot(Predict(DRselect, fun=plogis));plot(Predict(bDNselect, fun=plogis));plot(rms::calibrate(bDPNselect, method="boot", B=500), las=1);plot(rms::calibrate(bANselect, method="boot", B=500), las=1);plot(rms::calibrate(DRselect, method="boot", B=500), las=1);plot(rms::calibrate(bDNselect, method="boot", B=500), las=1); rms::validate(bDPNselect, method="boot", B=500);rms::validate(bANselect, method="boot", B=500);rms::validate(DRselect, method="boot", B=500);rms::validate(bDNselect, method="boot", B=500);plot(nom.dpn);plot(nom.an);plot(nom.dr);plot(nom.dn)



# cowplot::plot_grid(a,b,c,d,
#                    e,f,g,h,
#                    i,j,k,l,
#                    m,n,o,p,
#                    q,r,s,t,
#                    nrow = 5, ncol = 4, scale = .9, labels = "AUTO"); cowplot::plot_grid(a2,c2,
#                    e2,g2, 
#                    i2,k2,
#                    m2,o2,
#                    nrow = 4, ncol = 2, scale = .9, labels = "AUTO")


```



```{r extra rms functions}
#chunk test for multicollinearity
f <- ols(y ~ age + tricep + waist + pol(hip,2))
anova(f, tricep, waist, hip)  # 4 d.f. test

#penalization, ridge
pentrace(m4, seq(.01, .1, by = .01))
m5 <- update(m4, penalty = .01)
m5

#output exact function useing
(pred_logit <- Function(m4))

#nomogram
nom.dpn<-nomogram(DPNselect,
                 lp.at = seq(-3,4,by=0.5),
                 fun = function(x)1/(1+exp(-x)),
                     fun.at=c(0.001,0.01,.05,seq(0.1,0.9,by=0.1),0.95,0.99,0.999),
                     funlabel="Risk of Complication",
                 interact = list(Dur=c(5,10,20,30,40)),
                 conf.int=c(0.1,0.7),
                 abbrev=T,
                 minlength=1,lp=F)
```

```{r full vs select logreg models, realize goal of paper is not to make a subset model which matches full model, but to predict risk in complication naive individuals}
#DPN
DPNfull <- lrm(Peripheral_Neuropathy3 ~ Autonomic_Neuropathy3+
    DM_Retinopathy3+
    DM_Nephropathy3+
    
    Sex+
    Smoking+
    
    HTN3+
    Dyslipidemia3+

    Dr_Age+
    Dx_Age+
    Dur+
    
    Hemoglobin+
    Albumin+
    LDL+
    Total_Cholesterol+
    Triglycerides+
    HDL+
    avg.BUN+
    avg.Creatinine+
    avg.Systolic+
    avg.diastolic+
    avg.MicroAlb+
    ACR+
    AvgA1c+
    SDA1c+
    HbA1c_3.1+
    Max_HbA1c,
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)

DPNselect <- lrm(Peripheral_Neuropathy3 ~
    Autonomic_Neuropathy3+
    DM_Retinopathy3+
    DM_Nephropathy3+
    rcs(Dr_Age,3)+
    rcs(Dur,3)+
    rcs(Dx_Age,5)+
    rcs(avg.Systolic,3)+
    rcs(AvgA1c,4),
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)

lrtest(DPNfull,DPNselect)
DPNfull
DPNselect

#AN
ANfull <- lrm(Autonomic_Neuropathy3 ~ Peripheral_Neuropathy3+
    DM_Retinopathy3+
    DM_Nephropathy3+
    
    Sex+
    Smoking+
    
    HTN3+
    Dyslipidemia3+

    Dr_Age+
    Dx_Age+
    Dur+
    
    Hemoglobin+
    Albumin+
    LDL+
    Total_Cholesterol+
    Triglycerides+
    HDL+
    avg.BUN+
    avg.Creatinine+
    avg.Systolic+
    avg.diastolic+
    avg.MicroAlb+
    ACR+
    AvgA1c+
    SDA1c+
    HbA1c_3.1+
    Max_HbA1c,
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)

ANselect <- lrm(Autonomic_Neuropathy3 ~
    Peripheral_Neuropathy3+
    DM_Retinopathy3+
    DM_Nephropathy3+
    rcs(Dr_Age,3)+
    rcs(Dur,3)+
    rcs(Dx_Age,5)+
    rcs(avg.Systolic,3)+
    rcs(AvgA1c,4),
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)

lrtest(ANfull,ANselect)
ANfull
ANselect

#DR
DRfull <- lrm(DM_Retinopathy3 ~ Autonomic_Neuropathy3+
    Peripheral_Neuropathy3+
    DM_Nephropathy3+
    
    Sex+
    Smoking+
    
    HTN3+
    Dyslipidemia3+

    Dr_Age+
    Dx_Age+
    Dur+
    
    Hemoglobin+
    Albumin+
    LDL+
    Total_Cholesterol+
    Triglycerides+
    HDL+
    avg.BUN+
    avg.Creatinine+
    avg.Systolic+
    avg.diastolic+
    avg.MicroAlb+
    ACR+
    AvgA1c+
    SDA1c+
    HbA1c_3.1+
    Max_HbA1c,
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)

DRselect <- lrm(DM_Retinopathy3 ~
    Autonomic_Neuropathy3+
    Peripheral_Neuropathy3+
    DM_Nephropathy3+
    rcs(Dr_Age,3)+
    rcs(Dur,3)+
    rcs(Dx_Age,5)+
    rcs(avg.Systolic,3)+
    rcs(AvgA1c,4),
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)

lrtest(DRfull,DRselect)
DRfull
DRselect

#DN
DNfull <- lrm(DM_Nephropathy3 ~ Autonomic_Neuropathy3+
    Peripheral_Neuropathy3+
    DM_Retinopathy3+
    
    Sex+
    Smoking+
    
    HTN3+
    Dyslipidemia3+

    Dr_Age+
    Dx_Age+
    Dur+
    
    Hemoglobin+
    Albumin+
    LDL+
    Total_Cholesterol+
    Triglycerides+
    HDL+
    avg.BUN+
    avg.Creatinine+
    avg.Systolic+
    avg.diastolic+
    avg.MicroAlb+
    ACR+
    AvgA1c+
    SDA1c+
    HbA1c_3.1+
    Max_HbA1c,
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)

DNselect <- lrm(DM_Nephropathy3 ~
    Autonomic_Neuropathy3+
    Peripheral_Neuropathy3+
    DM_Retinopathy3+
    rcs(Dr_Age,3)+
    rcs(Dur,3)+
    rcs(Dx_Age,5)+
    rcs(avg.Systolic,3)+
    rcs(AvgA1c,4)+
    rcs(avg.MicroAlb,3),
          data = T1Dcomp_alldata, x = TRUE, y = TRUE)

lrtest(DNfull,DNselect)
DNfull
DNselect

```

Figure 2. Multivariate logistic regression models for 4 complications
A. Variables with coefficients
B. Nomogram
C. AUC of test data
D. Interation of Age and T1D Dur

#Comparing predictive models
```{r feature selection method comparison, not useful according to Harrell}

library(caret)
library(doParallel)
library(dplyr)
library(pROC)
library(glmnet)
source("my_functions.R")
T1Dcomp_alldata<-read.csv("2_T1DComplications data_cleaned for Paul.csv")

set.seed(123)



 # 1) DPN

dpn <- suppressWarnings(compare_3_models(T1Dcomp_alldata,"Peripheral_Neuropathy3",varstoassess = c("Peripheral_Neuropathy3","Autonomic_Neuropathy3","DM_Retinopathy3","DM_Nephropathy3","HTN3","Dyslipidemia3","CAD3","Prior_TIA3","Prior_MI3","Prior_CVA3","Dr_Age","Dur","Hemoglobin","Albumin","avg.Creatinine","avg.Systolic","avg.BUN","avg.MicroAlb","AvgA1c","Sex","HDL","Max_HbA1c")))

 # 2) DAN
can <- suppressWarnings(compare_3_models(T1Dcomp_alldata,"Autonomic_Neuropathy3",varstoassess = c("Autonomic_Neuropathy3","Peripheral_Neuropathy3","DM_Retinopathy3","DM_Nephropathy3","HTN3","Dyslipidemia3","CAD3","Prior_TIA3","Prior_MI3","Prior_CVA3","Dr_Age","Dur","Hemoglobin","Albumin","avg.Creatinine","avg.Systolic","avg.BUN","avg.MicroAlb","AvgA1c","Sex","HDL","Max_HbA1c")))

 # 3) Dret
dr <- suppressWarnings(compare_3_models(T1Dcomp_alldata,"DM_Retinopathy3",varstoassess = c("DM_Retinopathy3","Peripheral_Neuropathy3","Autonomic_Neuropathy3","DM_Nephropathy3","HTN3","Dyslipidemia3","CAD3","Prior_TIA3","Prior_MI3","Prior_CVA3","Dr_Age","Dur","Hemoglobin","Albumin","avg.Creatinine","avg.Systolic","avg.BUN","avg.MicroAlb","AvgA1c","Sex","HDL","Max_HbA1c")))

 # 4) Dnep
dn <- suppressWarnings(compare_3_models(T1Dcomp_alldata,"DM_Nephropathy3",varstoassess = c("DM_Nephropathy3","Peripheral_Neuropathy3","Autonomic_Neuropathy3","DM_Retinopathy3","HTN3","Dyslipidemia3","CAD3","Prior_TIA3","Prior_MI3","Prior_CVA3","Dr_Age","Dur","Hemoglobin","Albumin","avg.Creatinine","avg.Systolic","avg.BUN","avg.MicroAlb","AvgA1c","Sex","HDL","Max_HbA1c")))


#Summary data
paste("____________________________DPN____________________________");dpn$lasso_tst_roc$auc;dpn$rfelogit_tst_roc$auc;dpn$customlogit_tst_roc$auc;dpn$customlogit_tst_roc2$auc;dpn$lasso_coef;dpn$rfelogit_coef;dpn$customlogit_coef;dpn$customlogit_coef2;paste("____________________________CAN____________________________");can$lasso_tst_roc$auc;can$rfelogit_tst_roc$auc;can$customlogit_tst_roc$auc;can$customlogit_tst_roc2$auc;can$lasso_coef;can$rfelogit_coef;can$customlogit_coef;can$customlogit_coef2;paste("____________________________DR____________________________");dr$lasso_tst_roc$auc;dr$rfelogit_tst_roc$auc;dr$customlogit_tst_roc$auc;dr$customlogit_tst_roc2$auc;dr$lasso_coef;dr$rfelogit_coef;dr$customlogit_coef;dr$customlogit_coef2;paste("____________________________DN____________________________");dn$lasso_tst_roc$auc;dn$rfelogit_tst_roc$auc;dn$customlogit_tst_roc$auc;dn$customlogit_tst_roc2$auc;dn$lasso_coef;dn$rfelogit_coef;dn$customlogit_coef;dn$customlogit_coef2


```

Figure 3. Website developed to motivate change

Key features are risk reporting for 4 complications. How would risk be modified with changes in SBP or HbA1C? How patient compares to others of same Age and T1D dur for modifiable risk factors? Methods to modify risk factors (diet, weight loss association with BP). 

Conclusion: SBP and HbA1C are modifiable risk factors for T1D complications through lifestyle changes and medical management. Motivation to manage SBP and HbA1C may be more difficult for younger individuals and T1D complication risk tool may aid with this.